---
title: "SQ vs Lung Single Cell Analysis"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  tidy = TRUE,
  tidy.opts = list(width.cutoff = 120),
  message = FALSE,
  warning = FALSE
)
```

### Load In Myeloid Pre Processed Data 
Load Data Set 

```{r init}
library(dplyr)
library(Seurat)

```

```{r load}
# Load the PBMC dataset
cellData <- readRDS(file = "./mye3_w_alvmac_3_mitoThreshold_scaled_seuratv3.rds")

cellData
```

### Preform PCA on the data 

PCA on the scaled data. 

```{r pca,results='hide'}
cellData <- RunPCA(object = cellData, features = VariableFeatures(object = cellData))
```

Jack Straw plot - to see dimentions and cut off for PCs
```{r jackstraw, fig.height=6, fig.width=10}
# NOTE: This process can take a long time for big datasets, comment out for expediency. More approximate techniques such as those implemented in ElbowPlot() can be used to reduce computation time
cellData <- JackStraw(object = cellData, num.replicate = 100)
cellData <- ScoreJackStraw(object = cellData, dims = 1:20)
```


Elbow plot 

```{r elbow_plot, fig.height=6, fig.width=10}
ElbowPlot(object = cellData)
```


Heat Maps of PCs - can see which ones are clean and distinct 

```{r multi-heatmap, fig.height=12, fig.width=9}
DimHeatmap(object = cellData, dims = 1:15, cells = 500, balanced = TRUE)
```

### Cell Clustering 

CLuster them based on neighboors

```{r cluster, fig.height=5, fig.width=7}
cellData <- FindNeighbors(object = cellData, dims = 1:20)
cellData <- FindClusters(object = cellData, resolution = 0.4)

```


```{r tsne, fig.height=5, fig.width=7}
cellData <- RunUMAP(object = cellData, dims = 1:20)
```
Draw UMAP Plot 


```{r tsneplot, fig.height=5, fig.width=7}
# note that you can set `label = TRUE` or use the LabelClusters function to help label individual clusters
DimPlot(object = cellData, reduction = 'umap')
```
Save the UMAP plot to the file

```{r saveobject, eval=FALSE}
saveRDS(cellData, file = "./processed.rds")
```

### Labeling the Cell clusters 


Find the pos distinct markers for each cluster across the data 

```{r markers1, fig.height=8, fig.width=15}

cellData <- readRDS(file="./processed.rds")
cellData.markers <- FindAllMarkers(object = cellData, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
```

```{r markers2, fig.height=8, fig.width=15}
cellData.markers %>% group_by(cluster) %>% top_n(n = 15, wt = avg_logFC)

```
11 - DCs
SIGLECH
CCR9
ATP1B1
SELL
LY6D
SPIB
KLK1
SLA2
CD4
IGHM
TEX2
TCF4
RUNX2
BST2
IRF8

7 - Mac
TOP2A
ANLN
KIF23
RRM1
MARCO
KRT79
CTSK
KRT19
CHI3L3
CAR4
ATP6V0D2
MKI67
LPL
SORT1
TUBB5


10 - Dendritic Cells
ITGAE
CKB
SEPT3
CLEC9A
BTLA
FLT3
XCR1
ARSB
WDFY4
PLBD1
CST3
NAAA
IRF8
RAB43
RAB43

9 - Mac
GPNMB
PSAP
ANPEP
CTSD
CTSS
CD63
CTSB
PLD3
TREM2
HEXA
LGALS3
MYO5A
APOE
LGALS3BP
MFGE8

7 - Pre B or T cell 
TOP2A
MKI67
NCAPD2
KIF11
CCNA2
BIRC5
CENPE
CDK1
D17H6S56E-5
MCM5
RRM1
TUBB5
MCM6
MCM3
TMPO


8 - Mono Ly6c-
ACE
EMR4
SPN
CD300E
ITGAL
SLC12A2
TREML4
ENO3
ZFYVE9
CEACAM1
PGLYRP1
CYFIP2
STK10
DUSP16
NR4A1

0 - Mac AVL_LU or Gran 
CHI3L3
LPL
EAR2
CTSD
ATP6V0D2
ABCG1
CTSK
KRT19
KRT79
CD9
CAR4
F7
KLHDC4
DST
GDF15

6 - Stroma or DC
FSCN1
CCR7
CCL22
CACNB3
TMEM123
TBC1D4
CCL5
SERPINB9
SERPINB6B
IL4I1
RAMP3
SLC4A8
ADCY6
IL12B
ROGDI


5 - DC or Mac
CD209A
IL1R2
ITGAE
H2-EB1
H2-AB1
NAPSA
H2-DMB2
H2-AA
KMO
CCL17
MGL2
CST3
IFITM1
CDH1
MDH2


4 - Monocytes Ly6C+
HP
LY6C2
PLAC8
CD177
IFITM3
CD97
SELL
SORL1
MS4A4C
EMILIN2
ITGB7
SAMHD1
CCR2
ARL5C
RNF213

1 - Mac AT
CCL8
CD163
CBR2
CCL7
SEPP1
GAS6
FOLR2
C4B
STAB1
LYVE1
F13A1
FCNA
CCL2
FCGRT
TIMP2


12 - Mac AVL_LU
CHI3L3
EAR2
LPL
ATP6V0D2
ABCG1
KRT19
CD9
CAR4
KRT79
SORT1
EAR1
1600029D21RIK
SERPINE1
ADIPOR2
SIGLEC5

3 - Gran 
THBS1
PTGS2
VCAN
IL1B
CD14
SDC4
ITGAM
CCR2
TGFBI
VEGFA
ERRFI1
OSM
IL1RN
MMP19
FN1

2 - Gran 
MMP12
MMP9
MMP13
ARG1
CCL24
INHBA
TNFRSF9
ACP5
MMP14
ADAM8
PLA2G7
CLEC4N
VEGFA
SPP1
CXCL3

Macrophage - Cd14,Cd16,Cd64, CD68, Cd71, CCR5
CD64 - FCGR1
CD68 - CD68 
CD71 - TFRC

M1
CD68 - CD68
CD80 - CD80
CD86 - CD86
IL-1R - IL1R1
TLR2 - TLR2
TLR4 - TLR4
iNOS - NOS2
SOCS3

# 10, 12, 3, 
```{r markerplots mac, fig.height=8, fig.width=15}
#General Mac Markers
VlnPlot(object = cellData, features = c("FCGR1", "CD68", "TFRC", "CCR5"))
#M1 Markers 
VlnPlot(object = cellData, features = c("CD68", "CD80", "CD86", "IL1R1", "TLR2", "TLR4", "NOS2", "SOCS3"))

#FeaturePlot(cellData, reduction = "umap", features = c("FCGR3A", "MS4A7"))
```


M2: # 3 
CD115 - CSF1R
CD204 - MSR1
CD163
CD206 (MMR) - MRC1
CD209 (DC-SIGN) - CD209A
FcεR1 - FC
VSIG4
```{r markerplots mac169, fig.height=8, fig.width=15}
#169+
VlnPlot(object = cellData, features = c("SIGLEC1", "ITGAX"))

#M2 Macrophages 
VlnPlot(object = cellData, features = c("CSF1R", "MSR1","CD163","MRC1", "CD209A", "FC", "VSIG4"))
```
```{r markerplots mmm, fig.height=8, fig.width=15}
#Mature Mac Marker 
VlnPlot(object = cellData, features = c("MSR1"))

```

Monocyte - CD14+, CD16-/+ (two subsets) and LY6C lo and high subsets 
CD31 - PECAM1 
CD32 - FCGR2B
CD11b - ITGAM
CD43/Ly48 - SPN
CD45 - PTPRC
CD62L - SELL
CD115 - CSF1R
F4/80 - ADGRE1
GR1 - GSR
VEGF - vegfa

Monocyte Surface: 
CSF1R
LY6C

```{r markerplots mono1, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("CSF1R",  "LY6C2", "LY6A"))

```
LyC6+ - #4

```{r markerplots mono2, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("LY6C2","CCR1", "CCR2","ITGAM", "FCGR2B", "PECAM1", "SPN", "CD44", "PTPRC", "SELL", "CSF1R", "CX3CR1", "ADGRE1", "GSR", "VEGFA" ))

```

CMP - key marker CD123 - IL3RA
```{r markerplots CMP, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("IL3RA"))

```

Mast Cell 
CD117 - KIT
FCER1 - CDCA7L and RRM1
```{r markerplots Mast, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("KIT", "CDCA7L","RRM1"))

```


Neutrophil
CD66a - CEACAM1, CEACAM2
Ly6G
```{r markerplots Neu, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("CEACAM1", "CEACAM2", "LY6G"))

```


Basophil -
FCER1 - CDCA7L and RRM1
```{r markerplots Bas, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("CDCA7L", "RRM1"))

```
Eosphil
Cd170 - SIGLECF
CCR3/CD193
```{r markerplots Eos, fig.height=8, fig.width=15}
VlnPlot(object = cellData, features = c("SIGLECF", "CCR3"))

```

Names 
"Gran", "Gran, "Mac","Mac", "Ly6C+ Mono", "DC/Mac", "Mac/Stroma", "Mac/Gran", "LyC6- Mono", "Pre T/B", "Mac", "DC", "Mac", "DC"

Cd170 - SIGLECF
CCR3/CD193
```{r nameing, fig.height=8, fig.width=15}


new.cluster.ids <- c("Gran", "Mac", "Gran", "Gran", "Ly6C+ Mono", "DC/Mac", "Mac/Stroma", "Mac/Gran", "LyC6- Mono", "Mac", "DC", "DC", "Mac")
names(new.cluster.ids) <- levels(cellData)
cellData <- RenameIdents(cellData, new.cluster.ids)
DimPlot(cellData, reduction = "umap", label = TRUE, pt.size = 0.5)

```


```{r save, fig.height=8, fig.width=15}
saveRDS(cellData, file = "monocyteData.rds")

```

```{r UMAP}
DimPlot(cellData, reduction = "umap", label = TRUE, pt.size = 0.5)
```

## Counting Cells 
```{r count, fig.height=8, fig.width=15}

cellData <- readRDS(file = "./monocyteData.rds")


table(Idents(cellData))

table(Idents(cellData), cellData$loctreat)
```


```{r heatmap}
#cluster.averages <- AverageExpression(cellData, return.seurat = TRUE, add.ident = "loctreat")

DoHeatmap(cluster.averages, features = unlist(TopFeatures(cellData[["pca"]], balanced = TRUE)), size = 3, 
    draw.lines = FALSE)

cellData$CellType <- Idents(cellData)

```



```{r keycytokines}

Idents(cellData) <- "loctreat"
Idents(cellData)
cluster.averages <- AverageExpression(cellData, return.seurat = TRUE, add.ident = "CellType")

featuresCytokines = c("IL2", "IL12A", "IL12B","IL6")

DoHeatmap(cellData, features = unlist(featuresCytokines), size = 5, 
    draw.lines = TRUE, group.by = "CellType")

# T_L  T_Q UT_L UT_Q

```

#Inhibitory / Stim Markers on Monocytes 

PD1 - PDCD1
PDL1 - CD274
   

```{r loctreat}
#Idents(cellData) <- "loctreat"

#cluster.averages <- AverageExpression(cellData, return.seurat = TRUE)

features<- c("CCL2", "IFNG", "IL1A", "IL1B", "IL15", "IL18", "IFN", "CSF1", "IL12A", "IL12B")
badFeatures <- c("IL4", "IL13", "TGFB1", "TGFB2", "PTGER3", "IL6", "CCL5", "IL10", "HIF1A")
growthFactors <- c( "VEGFA", "MMP2", "MMP13", "MMP14", "MMP15", "MMP16", "MMP17")
#PGE2 = PTGER3

DotPlot(object = cluster.averages, features = badFeatures)
```

```{r counts of cells}
library(ggplot)

tableSub <- table(Idents(cellData), cellData$loctreat)


tabledf <- as.data.frame(tableSub)


barplot(tableSub, main="Counts",
   xlab="Condition", hue=col(tableSub),
  legend = rownames(tableSub), )

ggplot(data=tabledf, aes(x=Var2, y=Freq, fill=Var1)) +
geom_bar(stat="identity", position=position_dodge())

```


Pathway analysis
```{r}
rm(list=ls())
library(dplyr)
library(Seurat)
library(ggplot2)
library(cowplot)
library(gridExtra)    
library(EnhancedVolcano)
```


```{r}
seurat <- readRDS("./monocyteData.rds")

seurat <- subset(x = seurat, subset = loctreat == "T_Q" & "T_L")
unique(seurat$loctreat)
```


```{r}
# pathway scoring
library(fgsea)
library(GSA)
g <- rownames(GetAssayData(seurat, slot = "data"))

kegg = GSA.read.gmt("./biocarta.gmt")
```

```{r}
for (i in 1:length(kegg$geneset.names)) {
  set_name <- kegg$geneset.names[i]
  seurat = AddModuleScore(seurat, features = kegg[[1]][i], name = set_name)
}
```

```{r}
colnames(seurat@meta.data)
```

```{r}
kegg = GSA.read.gmt("./kegg.gmt")
```

```{r}
pvals =c()
dir =c()
for (i in 1:length(kegg$geneset.names)) {
  data = seurat@meta.data[,paste0(kegg$geneset.names[i],1)]
  result = wilcox.test(data[seurat$seurat_clusters == 0], data[seurat$seurat_clusters == 1])
  pvals = c(pvals, result$p.value)
 direction = ifelse(mean(data[seurat$seurat_clusters == 0]) > mean(data[seurat$seurat_clusters ==1]), 'Up', 'Down')
    dir = c(dir, direction)
}
df = data.frame(path = kegg$geneset.names, pval = pvals, direction = dir)
df = df[order(df$pval),]
#df$origin = 'kegg'
df$path = as.character(df$path)
df
write.csv(df, "biocarta.csv")

palette = c('#3e58a8', '#ee3424')

df$path = sapply(df$path, function(x) gsub(x= x, pattern = "KEGG_", replacement = ""))
plot_hallmarks <- ggplot(df[1:30,], aes(y = path, x = -log10(pval), fill = direction)) + 
  geom_col() + scale_y_discrete(limits = rev(df$path[1:30])) +
  scale_fill_manual(values = palette) +
  guides(fill = FALSE) + labs(y = c()) + 
  theme(panel.border = element_rect(colour = "black", fill=NA), panel.background = element_blank()) +
  ggtitle("L_v_Q_biocarta")

#pdf(file="zsgreen_tumor_cells_pdfs/pathway_biocarta.pdf", width = 6, height = 4)
plot_hallmarks
```




